---
title: "birds_BRT Manual"
subtitle: "v.`r SpaDES.core::moduleMetadata(module = 'birds_BRT', path = '..')$version`"
date: "Last updated: `r Sys.Date()`"
output:
  bookdown::html_document2:
    toc: true
    toc_float: true
    theme: sandstone
    number_sections: false
    df_print: paged
    keep_md: yes
editor_options:
  chunk_output_type: console
  bibliography: citations/references_birds_BRT.bib
link-citations: true
always_allow_html: true
---

# birds_BRT Module

<!-- the following are text references used in captions for LaTeX compatibility -->
(ref:birds-BRT) *birds_BRT*

```{r setup-birds-BRT, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  eval = TRUE,
  results = "hold"
)
```

#### Authors:

`r paste(as.character(SpaDES.core::moduleMetadata(module = "birds_BRT", path = '..')$authors), sep = ', ')`
<!-- ideally separate authors with new lines, '\n' not working -->

## Module Overview

### Module summary

<!-- TODO: update description for ON and other study areas -->
<!-- see https://borealbirds.github.io/methods -->

The birds module for loads an existing glm object for each species located in  
<https://drive.google.com/open?id=1obSvU4ml8xa8WMQhQprd6heRrN47buvI>. 
The model uses dynamic (vegetation: biomass of tree species) and static layers (Non-veg: WAT = water (1/0), URBAG = urban/agriculture (1/0), lLED25 = lake edge density with 5x5 moving window (continuous), DEV25 = development proportion within 5x5 moving window (continuous), and landform (categorical).

North American landcover 2005 (MODIS) is
source for all but landform (from AdaptWest land facet datset).
Vector ruggedness (already available - 
<https://drive.google.com/open?id=1dgIw70mDpDYrBExA52SkPoS1TFaZdLE9>) and road density (that should be available from the Anthropogenic module) will be added to the next version of these models.
The birds' prediction is masked to uplands as we do not have data for lowlands.

### Module inputs and parameters

Table \@ref(tab:moduleInputs-birds-BRT) shows the full list of module inputs.

```{r moduleInputs-birds-BRT, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE}
df_inputs <- SpaDES.core::moduleInputs("birds_BRT", "..")
knitr::kable(df_inputs,
             caption = "List of (ref:birds-BRT) input objects and their description.") |>
  kableExtra::kable_styling(latex_options = "scale_down", full_width = TRUE)
```

Provide a summary of user-visible parameters (Table \@ref(tab:moduleParams-birds-BRT))


```{r moduleParams-birds-BRT, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE}
df_params <- SpaDES.core::moduleParams("birds_BRT", "..")
knitr::kable(df_params, caption = "List of (ref:birds-BRT) parameters and their description.") |>
  kableExtra::kable_styling(latex_options = "scale_down", full_width = TRUE)
```

### Events

1. `init`;

2. `loadModels`;

3. `loadFixedLayers`;

4. `gettingData`;

5. `predictBirds`;

### Plotting

<!-- TODO: describe plot outputs -->

Write what is plotted.

### Saving

<!-- TODO: describe file outputs -->

Write what is saved.

### Module outputs

Description of the module outputs (Table \@ref(tab:moduleOutputs-birds-BRT)).

```{r moduleOutputs-birds-BRT, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE}
df_outputs <- SpaDES.core::moduleOutputs("birds_BRT", "..")
knitr::kable(df_outputs, caption = "List of (ref:birds-BRT) outputs and their description.") |>
  kableExtra::kable_styling(latex_options = "scale_down", full_width = TRUE)
```

### Links to other modules

Describe any anticipated linkages to other modules, such as modules that supply input data or do post-hoc analysis.

### Getting help

<https://github.com/FOR-CAST/birds_BRT/issues>
